\textbf{ Analytic Light Curves in Reflected Light: \\% Phase Curves, Occultations, and Non-Lambertian Scattering for Spherical Planets and Moons }

### 2020-05-19 tex/ms.tex
tiny tweaks

There are several limitations to the \starry model that are important
to bear in mind. The primary limitation concerns the maximum spherical
harmonic degree of the model. While the expressions derived here are
valid at arbitrary degree, we find that their numerical stability begins
to degrade above $l \sim 20-25$ for occultations and $l \sim 35-40$ for phase
curves. The same is true for the model in thermal light \citep{Luger2019},
and is due to (1) the large condition number of the change-of-basis matrix
from spherical harmonics to polynomials and (2) instabilities in the
many recursion relations used to evaluate the solution vectors
$\sT$ and $\rT$.
In principle, one could improve the numerical stability by evaluating
all expressions at higher floating-point precision, but in practice
the computational cost of this becomes quickly prohibitive.
However, it is important to keep in mind that our current best image
of an exoplanet is the $l=1$ map of HD189733b
\citep{Knutson2007,Majeau2012,deWit2012}. Next-generation facilities
such as JWST may allow us to probe surface modes as small as $l=5$
for some planets (Luger et al., \emph{in prep}), but even with future
telescopes such as LUVOIR it is extremely unlikely we will do better than
$l=20$. If we are ever able to probe resolutions smaller than
$180^\circ / 20 = 9^\circ$ on the surface, the \starry algorithm
will have to be revisited.

@@

The third limitation concerns the flexibility of the \starry model.
While we presented ways to capture non-Lambertian scattering
in \starry (\S\ref{sec:nonlambertian}), there are certain aspects
of light curves in reflected light that cannot be captured by the model.
One example of this is shadowing. Craters on the moon or
volcanoes on Io can cast large shadows visible from space, particularly
if viewed near crescent phase. Unfortunately, there is no way to model
this within the \starry framework. Another example is multiple scattering,
as in optically thick atmospheres, for which a proper radiative transfer model
must be used. There is also the case of specular reflection,
or ``glint'', which is a pronounced signal for the Earth due its
oceans \citep{Robinson2014} and on Titan due to
its hydrocarbon lakes \citep{Barnes2011}. In principle, glint could be
modeled in the same way as non-Lambertian sccattering, by constructing the
linear operator $\mathbf{I}$ in such a way as to downweight portions of the
projected disk where $\vartheta_i \neq \vartheta_r$. In practice, however,
if the size of the glint spot is small (which is typically the case),
an expansion at extremely high $l$ ($l \sim 360$ in the case of the Earth)
would be required, which would not work for the reasons above.
Instead, it may be possible to combine the \starry algorithm with the
formalism of \citet{Haggard2018}, who derived analytic expressions for
phase curves of delta function maps ($\delta$-map), to model glint.
As discussed in \citet{LustigYaeger2018}, glint mapping is an extremely
powerful way to not only map terrestrial planets but also to confirm
their habitability via the presence of an ocean.  Finally, the presence
of time-variable features, such as clouds, dust storms, or seasonal
variations of vegetation are not accounted for in the model.

>>>>

There are several limitations to the \starry model that are important
to bear in mind. The primary limitation concerns the maximum spherical
harmonic degree of the model. While the expressions derived here are
valid at arbitrary degree, we find that their numerical stability begins
to degrade above $l \sim 20-25$ for occultations and $l \sim 35-40$ for phase
curves. The same is true for the model in thermal light \citep{Luger2019},
and is due to (1) the large condition number of the change-of-basis matrix
from spherical harmonics to polynomials and (2) instabilities in the
many recursion relations used to evaluate the solution vectors
$\sT$ and $\rT$.
In principle, one could improve the numerical stability by evaluating
all expressions at higher floating-point precision, but in practice
the computational cost of this becomes quickly prohibitive.
However, it is important to keep in mind that our current best image
of an exoplanet is the $l=1$ map of HD189733b
\citep{Knutson2007,Majeau2012,deWit2012}. Next-generation facilities
such as JWST may allow us to probe surface modes as small as $l=5$
for some planets (Luger et al., \emph{in prep}), but even with future
telescopes such as LUVOIR it is extremely unlikely we will do better than
$l=20$. If cases arise requiring a resolution smaller than about
$180^\circ / 20 = 9^\circ$ on the surface, the \starry algorithm
will have to be revisited.

@@

The third limitation concerns the flexibility of the \starry model.
While we presented ways to capture non-Lambertian scattering
in \starry (\S\ref{sec:nonlambertian}), there are certain aspects
of light curves in reflected light that cannot be captured by the model.
One example of this is shadowing. Craters on the moon or
volcanoes on Io can cast large shadows visible from space, particularly
if viewed near crescent phase. Unfortunately, there is no way to model
this within the \starry framework. Another example is multiple scattering,
as in optically thick atmospheres, for which a proper radiative transfer model
must be used. There is also the case of specular reflection,
or ``glint'', which is a pronounced signal for the Earth due its
oceans \citep{Robinson2014} and on Titan due to
its hydrocarbon lakes \citep{Barnes2011}. In principle, glint could be
modeled in the same way as non-Lambertian sccattering, by constructing the
linear operator $\mathbf{I}$ in such a way as to downweight portions of the
projected disk where $\vartheta_i \neq \vartheta_r$. In practice, however,
if the size of the glint spot is small (which is typically the case),
an expansion at extremely high $l$ ($l \sim 360$ in the case of the Earth)
would be required, which would not work for the reasons above.
Instead, it may be possible to combine the \starry algorithm with the
formalism of \citet{Haggard2018}, who derived analytic expressions for
phase curves of delta function maps ($\delta$-map), to model glint.
As discussed in \citet{LustigYaeger2018}, glint mapping is an extremely
powerful way to not only map terrestrial planets but also to confirm
their habitability via the presence of an ocean.  Lastly, the presence
of time-variable features, such as clouds, dust storms, or seasonal
variations of vegetation are not accounted for in the model, although
the \href{https://rodluger.github.io/starry}{documentation} discusses
how one may approach the modeling of temporal features.
